### 2020-04-11 Thesis-Chapter2-Background.tex
Move EA alternatives to own section
Work on intro, background

\newcommand{\lawstart}[0]{\begin{itemize}}
\newcommand{\law}[3]{\item #1: \textbf{#2} \cite{#3}} % Year, name, citation
\newcommand{\lawsend}{\end{itemize}}

\lawstart
    \law{1789}{\acf{AWA}}{congress_1789}
    \law{1791}{Bill of Rights: Amendments I, IV, and V}{madison_1791}
    \law{1976}{\acf{AECA}}{morgan_hr13680_1976}
% Established export controls over strong encryption techniques introduced ITAR (International Traffic in Arms
% Regulations) \cite{kehl_right_2015}
    \law{1978}{\acf{FISA}}{rodino_1978}
% (FISA has been the subject of frequent FBI abuse: \cite{shamsi_2011} \cite{tucker_2020})
% (the FBI almost lost some provisions; follow up on https://apnews.com/110ff653e98478da66765897dde3d613,
%     https://www.washingtontimes.com/news/2020/mar/15/adam-schiff-fisa-reform-intervention-protects-fbi-/,
%     https://saraacarter.com/fisa-senate-passes-measure-extending-surveillance-powers-abuses-not-addressed/)
    \law{1986}{\acf{CFAA}}{hughes_hr4718_1986}
% \cite{wolff_computer_2016}
    \law{1986}{\acf{ECPA}}{kastenmeier_hr4952_1986}
    \law{1994}{\acf{CALEA}}{edwards_hr4922_1994}
    \law{2001}{\acf{USAPATRIOT} Act}{sensenbrenner_2001}
    \law{2007}{\acf{PAA}}{mcconnell_s1927_2007}
    \law{2008}{\acf{FISAAA}}{reyes_hr6304_2008}
    \law{2015}{\acf{USAFREEDOM} Act}{sensenbrenner_2015}
\lawsend

@@

\lawstart
    \law{1993}{The \ac{clipperchip}}{press_1993}
% \cite{thompson_2015}
    \law{1996}{The SAFE Act}{goodlatte_hr3011_1996}
    \law{1997}{Encrypted Communications Privacy Act}{leahy_s376_1997}
% mix of SE protection and EA guidelines?
    \law{1998}{Encryption Protects the Rights of Individuals from Violation and Abuse in CYberspace (E-PRIVACY) Act}{ashcroft_s2067_1998}
% like above?
    \law{1999}{Encryption for the National Interest Act}{goss_hr2616_1999}
    \law{1999}{Tax Relief for Responsible Encryption Act}{goss_hr2617_1999}
% ^^ these two introduced by same rep on same day
    \law{2016}{Compliance with Court Orders Act (discussion draft)}{burr_2016}
    \law{2018}{The Secure Data Act}{lofgren_hr5823_2018}
% federal SE protections
    \law{2019}{The ENCRYPT Act}{lieu_hr4170_2019}
% state SE protections
    \law{2020}{The EARN IT Act}{graham_s3398_2020}
\lawsend

@@

\newcommand{\countrystart}[0]{\begin{itemize}}
\newcommand{\country}[1]{\item \textbf{#1:}} % Country
\newcommand{\countryend}{\end{itemize}}

\countrystart
    \country{Australia} In 2018 passed its Assistance and Access bill \cite{ag_2018} \cite{newman_2018}
    \country{UK} Investigatory Powers Act \cite{legislature_2016}
    \country{EU} GDPR \cite{parliament_2016}
    \country{China} See \cite{donahue_comparative_2018}
    \country{Israel} See \cite{donahue_comparative_2018}
    \country{Japan, Korea} can't find anything
    \country{General} See \cite{budish_encryption_2018}
\countryend

\section{Technical Approaches}
\label{sec-tech-approaches}

@@

\newcommand{\propstart}[0]{\begin{itemize}}
\newcommand{\prop}[2]{\item \textbf{#1} \cite{#2}} % Proposal name, citation
\newcommand{\propend}{\end{itemize}}

\propstart
    \prop{Clipper chip/third-party key escrow}{press_1993}
% toasted by \cite{blaze_protocol_1994}
    \prop{Oblivious key escrow}{goos_oblivious_1996}
    \prop{Partial key escrow}{denning_taxonomy_1996}
\propend

@@

\propstart
    \prop{CLEAR}{ozzie_2018}
    \prop{Device escrow with HW}{savage_lawful_2018}
    \prop{Computational, blockchain approaches}{phan_key_2017}
    \prop{Lawful hacking}{nguyen_lawful_2017}
% see also \cite{soesanto_2018} \cite{kerr_encryption_2017}
    \prop{Forced password disclosure}{kerr_encryption_2017}
% see also \cite{bittenbender_2019}
    \prop{Alternative paths to plaintext}{kerr_encryption_2017}
% access device while unlocked; find copies elsewhere
\propend

>>>>

\newcommand{\lawsstart}{\begin{itemize}}
\newcommand{\law}[4]{ % Year, name, citation, description
    \item #1: \textbf{#2} \cite{#3} \newline
    \parbox{\linewidth}{#4} \newline
}
\newcommand{\lawsend}{\end{itemize}}

\lawsstart
    \law{1789}{\acf{AWA}}{congress_1789}{}

    \law{1791}{Bill of Rights: Amendments I, IV, and V}{madison_1791}{}

    \law{1976}{\acf{AECA}}{morgan_hr13680_1976}{

Established export controls over strong encryption techniques introduced ITAR (International Traffic in Arms
Regulations) \cite{kehl_right_2015}

    }

    \law{1978}{\acf{FISA}}{rodino_1978}{}

% (FISA has been the subject of frequent FBI abuse: \cite{shamsi_2011} \cite{tucker_2020})
% (the FBI almost lost some provisions; follow up on https://apnews.com/110ff653e98478da66765897dde3d613,
%     https://www.washingtontimes.com/news/2020/mar/15/adam-schiff-fisa-reform-intervention-protects-fbi-/,
%     https://saraacarter.com/fisa-senate-passes-measure-extending-surveillance-powers-abuses-not-addressed/)

    \law{1986}{\acf{CFAA}}{hughes_hr4718_1986}{}

% \cite{wolff_computer_2016}

    \law{1986}{\acf{ECPA}}{kastenmeier_hr4952_1986}{}

    \law{1994}{\acf{CALEA}}{edwards_hr4922_1994}{}

    \law{2001}{\acf{USAPATRIOT} Act}{sensenbrenner_2001}{}

    \law{2007}{\acf{PAA}}{mcconnell_s1927_2007}{}

    \law{2008}{\acf{FISAAA}}{reyes_hr6304_2008}{}

    \law{2015}{\acf{USAFREEDOM} Act}{sensenbrenner_2015}{}

\lawsend

@@

\lawsstart
    \law{1993}{The \ac{clipperchip}}{press_1993}{}
% \cite{thompson_2015}
    \law{1996}{The SAFE Act}{goodlatte_hr3011_1996}{}
    \law{1997}{Encrypted Communications Privacy Act}{leahy_s376_1997}{}
% mix of SE protection and EA guidelines?
    \law{1998}{Encryption Protects the Rights of Individuals from Violation and Abuse in CYberspace (E-PRIVACY) Act}{ashcroft_s2067_1998}{}
% like above?
    \law{1999}{Encryption for the National Interest Act}{goss_hr2616_1999}{}
    \law{1999}{Tax Relief for Responsible Encryption Act}{goss_hr2617_1999}{}
% ^^ these two introduced by same rep on same day
    \law{2016}{Compliance with Court Orders Act (discussion draft)}{burr_2016}{}
    \law{2018}{The Secure Data Act}{lofgren_hr5823_2018}{}
% federal SE protections
    \law{2019}{The ENCRYPT Act}{lieu_hr4170_2019}{}
% state SE protections
    \law{2020}{The EARN IT Act}{graham_s3398_2020}{}
\lawsend

@@

\newcommand{\countriesstart}{\begin{itemize}}
\newcommand{\country}[1]{\item \textbf{#1:}} % Country
\newcommand{\countriesend}{\end{itemize}}

\countriesstart
    \country{Australia} In 2018 passed its Assistance and Access bill \cite{ag_2018} \cite{newman_2018}
    \country{UK} Investigatory Powers Act \cite{legislature_2016}
    \country{EU} GDPR \cite{parliament_2016}
    \country{China} See \cite{donahue_comparative_2018}
    \country{Israel} See \cite{donahue_comparative_2018}
    \country{Japan, Korea} can't find anything
    \country{General} See \cite{budish_encryption_2018}
\countriesend

\section{Approaches to Exceptional Access}
\label{sec-tech-approaches}

% There are several types of approaches:
% - Allow strong encryption (anyone with the users' private keys can see it)
% - Disallow encryption (anyone can see it)
% - Make encryption sufficiently weak, e.g. limit key length (anyone with enough CPU can see it)
% - Require use of backdoored algorithms, security via obscurity (anyone with the secret knowledge can see it)
% - Allow any encryption, just be able to provide keys (anyone with the extra keys can see it)
%   - Store extra in gov't DBs, in corporate DBs, in third-party DBs, on blockchain, in secure HW on device
%   - Split the key and store it multiple places
% - Use some alternative path
%   - Coerced passwords, finding unencrypted copies, seizing devices while unlocked, lawful hacking

@@

\newcommand{\propsstart}[0]{\begin{itemize}}
\newcommand{\prop}[2]{\item \textbf{#1} \cite{#2}} % Proposal name, citation
\newcommand{\propsend}{\end{itemize}}

\propsstart
    \prop{Clipper chip/third-party key escrow}{press_1993}
% toasted by \cite{blaze_protocol_1994}
    \prop{Oblivious key escrow}{goos_oblivious_1996}
    \prop{Partial key escrow}{denning_taxonomy_1996}
\propsend

@@

\propsstart
    \prop{CLEAR}{ozzie_2018}
    \prop{Device escrow with HW}{savage_lawful_2018}
    \prop{Computational, blockchain approaches}{phan_key_2017}
\propsend

\subsection{Alternative Approaches}

The following approaches do not use \ac{EA} at all, but are alternative technical or legal mechanisms that can be used
towards the same goal.

\propsstart
    \prop{Lawful hacking}{nguyen_lawful_2017}
% see also \cite{soesanto_2018} \cite{kerr_encryption_2017}
    \prop{Forced password disclosure}{kerr_encryption_2017}
% see also \cite{bittenbender_2019}
    \prop{Alternative paths to plaintext}{kerr_encryption_2017}
% access device while unlocked; find copies elsewhere
\propsend
